Generality in food webs scales with species richness, not latitude

### 2015-05-19 manuscript/latitudepaper.tex
This should be all the refs

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
  Food webs --networks of feeding links between species-- have been used for several decades to summarize the structure of 
  ecological communities~\citep{Petchey2008,Williams2000,Paine1980} and to understand how that structure relates to environmental variables 
  such as habitat type~\citep{Shurin2006,Briand1983}, primary productivity~\citep{Vermaat2009,Thompson2005a,Townsend1998}, and climate~\citep{Baiser2012,Petchey2010}. The latter
  variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
  climate is more variable at high latitudes~\citep{mappaper}. These variables affect both the resources available and species' 
  metabolisms~\citep{,Hechinger2011}, and have been proposed as determinants of the strength of 
  interspecific interactions~\citep{Lang2012,Schleuning2012,Schemske2009}. 
  By modulating interactions between species, latitudinal gradients may also shape food-web structure~\citep{}.
  Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
  latitudinal gradient in species richness, one of the most general and robust patterns in ecology~\citep{Schemske2009a,Stegen2009,Macpherson2002,Kaufman1995}.

@@

  One proposed link between species richness and latitude is the ``latitude-niche breadth hypothesis''~\citep{,Vazquez2004}. 
  This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
  turn should evolve smaller niches \citep{Vazquez2004}. These narrow niches should therefore allow more species to 
  coexist in the tropics than at higher latitudes. Alternatively, the higher productivity of the tropics \citep{Brown2004}
  may result in a broader niche space \citep{Davies2007} which could also sustain greater biodiversity even if niche 
  sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocally 
  supported \citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
  richness \citep{Lappalainen2006,Krasnov2008,Slove2010}.

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialization (i.e., niche breadth) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialization in the tropics~\citep{}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
  relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
  the ``latitude-niche breadth hypothesis''.

@@

  We compiled a list of 163 empirical food webs from
  multiple sources (see Supplemental Information for web origins and selection
  criteria). We recorded study site latitude from the original source where
  possible or, where study sites were described but exact coordinates were not
  given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.
  We then divided the species in each web into basal resources (those species with
  consumers but no prey), top predators (those species with prey but no predators),
  and intermediate consumers (species with both predators and prey, including cannibalistic
  species).

@@

  As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
  is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
  Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
  sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
  is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
  species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
  of the dataset; in each case using the number of species ($S$) and 
  links ($L$) in each web to calculate the mean number of links per species ($Z$), mean generality 
  ($G$), and mean vulnerability ($V$) of the web. 
  The version of the dataset used did not qualitatively change the results, suggesting that
  the scaling relationships between species richness, other food-web properties, and latitude are very 
  similar whether or not redundant species are included. For simplicity, we present only
  the results for the original webs.

@@

  Next, we examined the form of the scaling relationship between each 
  property (links per species, generality, and vulnerability) and 
  species richness. The scaling relationship between links per species ($Z$) and 
  species richness ($S$) has been shown to be a power law \citep{Riede2010} of the form

@@

  \noindent Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
  Specifically, equation~(\ref{Power}) implies a normally-distributed, additive error and equation~(\ref{Loglog}) a lognormal,
  multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
  over another, we follow the recommendations in \citet{Xiao2011} and compared the two
  model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
  data was then used to test for potential effects of latitude.

  Although scaling relationships between species richness and generality or
  species richness and vulnerability have not been explicitly examined (but see scaling 
  relationships for the standard devaitions of each property in \citet{Riede2010}), we expect that they will follow
  power laws similar to that of the relationship between species richness and links per species.
  This is because the links taken into account in calculating generality and vulnerability are subsets 
  of the total links included when calculating links per species. As with links per species, we explicitly 
  compared the error distributions of models for generality and vulnerability using
  both the power-law and logarithmic formulations. 
  In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
  species richness.

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\citep{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\citep{}, intuitively one might suspect that any relationship
  between food-web properties such as generality might be due to the latitudinal
  gradient in species richness. In this dataset, however, we did not find
  overall latitudinal gradients in species richness, links per species, 
  generality, vulnerability, or the proportions of food webs accounted for by 
  basal resources, intermediate consumers, and top predators.

  The lack of latitudinal gradients in food-web properties in this dataset contrasts
  strongly with other studie~\citep{}. As numbers of species and links included in a food web
  vary strongly with sampling effort as well as with the underlying diversity of the study
  area, it is possible that the lack of latitudinal trends here is a result of researchers
  tending to expend similar amounts of sampling effort across studies. This could result in
  food webs describing species-rich tropical communities omitting more species and links
  than species-poor arctic communities. In addition, it is worth noting that gradients in
  species richness are generally measured for a single taxonomic group at a time~\citep{}.
  It is possible that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in richness. In
  either case, the lack of association between species richness
  and latitude in any ecosystem type nevertheless means that any effect of latitude on scaling relationships
  between species richness and other properties is not being driven by underlying variation in
  species richness, allowing us to more clearly assess effects of latitude on scaling with 
  species richness and proportions of species in different trophic levels.

  Scaling of links per species, generality, and vulnerability with species
  richness varied strongly across ecosystem types. In estuarine, marine, and
  terrestrial food webs scaling of each property varied little with latitude.
  This is consistent with the idea that species' niche breadths do not vary
  systematically with temperature and productivity but that the niche space
  might be larger in species-rich communities~\citep{Davies2007}. Rather than
  niche space depending on temperature and productivity, it may be that species
  diversity itself affects the niche space available to species (although climate
  may affect speciation rates and therefore the diversity in a region~\citep()).

  Unlike other ecosystem types, in lake and stream food webs 
  the scaling of generality was stronger in higher-latitude food webs. 
  In lake food webs, this trend was echoed in the scaling relationships
  between species richness and links per species and vulnerability.
  This means that species in tropical freshwater
  communities gain fewer additional feeding links per additional species in the
  web and that species in tropical lakes also gain fewer predators, and fewer links
  in general, per additional species. These trends are consistent with the hypothesis
  that greater stability in the tropics leads to narrower niches~\citep{Brown2004} and a higher proportion of
  specialists, as has been observed in mutualist networks~\citep{}. Inhabitants of high-latitude
  freshwater communities may also be more generalist due to prey switching if they rely on a series
  of ephemeral prey such as maturing insect larvae~\citep{}.

>>>>

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
  Food webs --networks of feeding links between species-- have been used for several decades to summarize the structure of 
  ecological communities~\citep{Petchey2008,Williams2000,Paine1980} and to understand how that structure relates to environmental variables 
  such as habitat type~\citep{Shurin2006,Briand1983}, primary productivity~\citep{Vermaat2009,Thompson2005a,Townsend1998}, and climate~\citep{Baiser2012,Petchey2010}. The latter
  variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
  climate is more variable at high latitudes~\citep{Field1998}. These variables affect both the resources available and species' 
  metabolisms~\citep{Hechinger2011,OConnor2009,White2009,}, and have been proposed as determinants of the strength of 
  interspecific interactions~\citep{Lang2012,Schleuning2012,Schemske2009}. 
  By modulating interactions between species, latitudinal gradients may also shape food-web structure~\citep{}.
  Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
  latitudinal gradient in species richness, one of the most general and robust patterns in ecology~\citep{Schemske2009a,Stegen2009,Macpherson2002,Kaufman1995}.

@@

  One proposed link between species richness and latitude is the ``latitude-niche breadth hypothesis''~\citep{Vazquez2004}. 
  This hypothesis predicts that decreased seasonality in the tropics should lead to more stable populations, which in
  turn should evolve smaller niches~\citep{Vazquez2004}. These narrow niches should therefore allow more species to 
  coexist in the tropics than at higher latitudes. Alternatively, the higher productivity of the tropics~\citep{Brown2004}
  may result in a broader niche space~\citep{Davies2007} which could also sustain greater biodiversity even if niche 
  sizes are globally similar. Although the assumptions of the latitude-niche breadth hypothesis are only equivocally 
  supported~\citep{Vazquez2004}, it remains a compelling potential mechanism for the latitudinal gradient in species 
  richness~\citep{Lappalainen2006,Krasnov2008,Slove2010}.

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialization (i.e., niche breadth) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialization in the tropics~\citep{Dyer2007,Marra1997} but see~\citep{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialization may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialization by examining the shape of the scaling 
  relationship between specialization and species richness over changing latitude. Here, we use three measures of specialization:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialization implied by
  the ``latitude-niche breadth hypothesis''.

@@

  We compiled a list of 163 empirical food webs from
  multiple sources (see Supplemental Information for web origins and selection
  criteria). We recorded study site latitude from the original source where
  possible or, where study sites were described but exact coordinates were not
  given, obtained estimated coordinates using Google Earth~\citep{GoogleEarth}.
  We then divided the species in each web into basal resources (those species with
  consumers but no prey), top predators (those species with prey but no predators),
  and intermediate consumers (species with both predators and prey, including cannibalistic
  species).

@@

  As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
  is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
  Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
  sets to form ``trophic species'' webs (e.g.~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
  is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
  species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
  of the dataset; in each case using the number of species ($S$) and 
  links ($L$) in each web to calculate the mean number of links per species ($Z$), mean generality 
  ($G$), and mean vulnerability ($V$) of the web. 
  The version of the dataset used did not qualitatively change the results, suggesting that
  the scaling relationships between species richness, other food-web properties, and latitude are very 
  similar whether or not redundant species are included. For simplicity, we present only
  the results for the original webs.

@@

  Next, we examined the form of the scaling relationship between each 
  property (links per species, generality, and vulnerability) and 
  species richness. The scaling relationship between links per species ($Z$) and 
  species richness ($S$) has been shown to be a power law~\citep{Riede2010} of the form

@@

  \noindent Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
  Specifically, equation~(\ref{Power}) implies a normally-distributed, additive error and equation~(\ref{Loglog}) a lognormal,
  multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
  over another, we follow the recommendations in~\citet{Xiao2011} and compared the two
  model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
  data was then used to test for potential effects of latitude.

  Although scaling relationships between species richness and generality or
  species richness and vulnerability have not been explicitly examined (but see scaling 
  relationships for the standard devaitions of each property in~\citet{Riede2010}), we expect that they will follow
  power laws similar to that of the relationship between species richness and links per species.
  This is because the links taken into account in calculating generality and vulnerability are subsets 
  of the total links included when calculating links per species. As with links per species, we explicitly 
  compared the error distributions of models for generality and vulnerability using
  both the power-law and logarithmic formulations. 
  In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
  species richness.

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\citep{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\citep{Schemske2009a,Stegen2009,Macpherson2002,Kaufman1995}, intuitively one might suspect that any relationship
  between food-web properties such as generality might be due to the latitudinal
  gradient in species richness. In this dataset, however, we did not find
  overall latitudinal gradients in species richness, links per species, 
  generality, vulnerability, or the proportions of food webs accounted for by 
  basal resources, intermediate consumers, and top predators.

  The lack of a latitudinal gradient in species richness in this dataset contrasts
  strongly with other studies~\citep{Schemske2009a,Stegen2009,Macpherson2002,Kaufman1995}. As numbers of species and links included in a food web
  vary strongly with sampling effort as well as with the underlying diversity of the study
  area, it is possible that the lack of latitudinal trends here is a result of researchers
  tending to expend similar amounts of sampling effort across studies. This could result in
  food webs describing species-rich tropical communities omitting more species and links
  than species-poor arctic communities. In addition, it is worth noting that gradients in
  species richness are generally measured for a single taxonomic group at a time (e.g.~\citep{Kaufman1995}).
  It is possible that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in richness. In
  either case, the lack of association between species richness
  and latitude in any ecosystem type nevertheless means that any effect of latitude on scaling relationships
  between species richness and other properties is not being driven by underlying variation in
  species richness, allowing us to more clearly assess effects of latitude on scaling with 
  species richness and proportions of species in different trophic levels.

  Scaling of links per species, generality, and vulnerability with species
  richness varied strongly across ecosystem types. In estuarine, marine, and
  terrestrial food webs scaling of each property varied little with latitude.
  This is consistent with the idea that species' niche breadths do not vary
  systematically with temperature and productivity but that the niche space
  might be larger in species-rich communities~\citep{Davies2007}. Rather than
  niche space depending on temperature and productivity, it may be that species
  diversity itself affects the niche space available to species (although climate
  may affect speciation rates and therefore the diversity in a region~\citep{Currie2004}).

  Unlike other ecosystem types, in lake and stream food webs 
  the scaling of generality was stronger in higher-latitude food webs. 
  In lake food webs, this trend was echoed in the scaling relationships
  between species richness and links per species and vulnerability.
  This means that species in tropical freshwater
  communities gain fewer additional feeding links per additional species in the
  web and that species in tropical lakes also gain fewer predators, and fewer links
  in general, per additional species. These trends are consistent with the hypothesis
  that greater stability in the tropics leads to narrower niches~\citep{Brown2004} and a higher proportion of
  specialists. This may be partly due to high-latitude species
  tending to switch between different seasonally-available
  prey~\citep{Isaac2012,Wilhelm1999} while tropical freshwater ecosystems may have more stable composition.
